---
title: "Plot Instructions"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read Data

Data is kept one level above git repo. It is a very large file, so not included in Git repo.

```{r}
hspower <- read.table("../household_power_consumption.txt",header = TRUE, sep = ";")
```

```{r}
str(hspower)
```

There are 2075259 entries, which is a large dataframe. We need to make the following changes

-   Convert Date to a readable format

```{r}
library(lubridate)
## We will use lubridate functions
## Convert all columns 3 to 9 to numeric
hspower$Date <- dmy(hspower$Date)
```

```{r}
str(hspower) ## Check up
```

OK, we need only data from 2007-02-01 to 2007-02-02

```{r}
from <- ymd("2007-02-01")
to <- ymd("2007-02-02")
hspower <- hspower[hspower$Date %in% from:to,]
```

```{r}
dim(hspower) ## Checkup
```

Now there are 2880 values, down from 2075259. Easier and faster to handle. Now we need to

1.  Make columns 3 to 9 numeric

2.  Make a datetime column to combine Date and Time values

```{r}
hspower[,3:9]<- lapply(hspower[,3:9],as.numeric)
hspower$datetime <- ymd_hms(paste(hspower[,1],hspower[,2]))
```

```{r}
str(hspower) ## Checkup
```

OK 
## Plotting

### Plot 1

```{r}
hist(hspower$Global_active_power, col = "red", xlab = "Global Active Power (kilowatts)", main = "Global Active Power")
```

### Plot 2

```{r}
plot(hspower$datetime,hspower$Global_active_power, type = "l", xlab="",ylab="Global Active Power (kilowatts)")
```

### Plot 3

```{r}
plot(hspower$datetime,hspower$Sub_metering_1, type = "l", xlab="",ylab="Energy sub metering")
lines(hspower$datetime,hspower$Sub_metering_2,col = "red")
lines(hspower$datetime,hspower$Sub_metering_3,col = "blue")
legend(x="topright",legend=c("Sub_metering_1","Sub_metering_2","Sub_metering_3"), col=c("black","red","blue"), lwd=2, box.lwd=0)
```

### Plot 4

This requires generating two plots we have not generated before, and two (Plot 2 and 3) which we already have.

#### Plot 4 top right

```{r}
plot(hspower$datetime,hspower$Voltage, type = "l", xlab="datetime",ylab="Voltage")
```

#### Plot 4 bottom right

```{r}
plot(hspower$datetime,hspower$Global_reactive_power, type = "l", xlab="datetime",ylab="Global_Reactive_Power")
```

To get Plot 4, we need to plot these separately.The code for the individual plot functions is almost the same as the individual codes abobe, but some trial and error adjustments have been made.

```{r}
par(mfrow = c(2,2), mar=c(4,4,0,2))

plot(hspower$datetime,hspower$Global_active_power, type = "l", xlab="",ylab="Global Active Power")

plot(hspower$datetime,hspower$Voltage, type = "l", xlab="datetime",ylab="Voltage")

plot(hspower$datetime,hspower$Sub_metering_1, type = "l", xlab="",ylab="Energy sub metering")
lines(hspower$datetime,hspower$Sub_metering_2,col = "red")
lines(hspower$datetime,hspower$Sub_metering_3,col = "blue")
legend(x="topright",legend=c("Sub_metering_1","Sub_metering_2","Sub_metering_3"), col=c("black","red","blue"), lwd=2, box.lwd = 0)

plot(hspower$datetime,hspower$Global_reactive_power, type = "l", xlab="datetime",ylab="Global_Reactive_Power")

```

Great, this all works. But [our question](README.md) requires specific files. These files are obtained by 

1.    Generating the R source code from this document:
```{r eval = FALSE}
knitr::purl("CodeBook.Rmd",output="R_plots.R")
```
2.   Write R code to read the raw data and write filtered data as `data.txt`
```{r eval=FALSE}
library(lubridate)
hspower<-read.table("../household_power_consumption.txt",
                    header=TRUE,
                    sep=";"
         )
from <- ymd("2007-02-01")
to <- ymd("2007-02-02")
hspower <- hspower[dmy(hspower$Date) %in% from:to,]
write.table(hspower,col.names=TRUE,
            row.names=FALSE,sep=";",
            quote=FALSE,
            file="data.txt"
)
```

3.   Creating individual files for each plot. This is copied from `R_plots.R`. Keep data reading file separate, so it can be sourced for each plot, the Data_read.R file. 
```{r eval = FALSE}
hspower<-read.table("data.txt",header=TRUE,sep=";")
hspower[,3:9]<- lapply(hspower[,3:9],as.numeric)
hspower$datetime <- dmy_hms(paste(hspower[,1],hspower[,2]))
```

4.    Write the png generation code. At the beginning add
```{r eval = FALSE}
png("plot1.png", width = 480, height = 480, bg = "transparent")
```
Finish off with dev.off()
```{r eval = FALSE}
dev.off()
```

